FFT based convolution

### 2010-03-10 article/Article.tex
update article and style with the new IJ style
Ignore-this: 83f17015111d18b0752b208ac8ce75e9

darcs-hash:20100309042535-2fc9d-ead501d6d8029574558b94b7e7af0de3dd9982e6.gz
article continued
Ignore-this: 6b8dfb502f3001b366ff081bda9cfb0c

darcs-hash:20100310000931-2fc9d-9873e486bf49c07fcf8f97391d31823a6eb71e05.gz

@@

\begin{document}
\maketitle

@@

@@

\subsubsection{Still to be done}

There are still some missing features in the FFTW filters. The ones which comes to mind are:
\begin{itemize}
 \item let the user create a plan with \verb$FFTW_PATIENT$, \verb$FFTW_MEASURE$, or
\verb$FFTW_EXHAUSTIVE$, so that FFTW can create better plan with \verb$FFTW_ESTIMATE$ in a
later run;
 \item provide an API to save the FFTW wisdoms.
\end{itemize}

@@

@@

@@

Most of the time, and despite the usual advice to use a size which is a power of two,
those sizes perform significantly better than the closest sizes which are a power of two, probably
because of the smaller impact of the number of pixels in the image.

@@

@@

Some other work may be done on this basis -- an non exhaustive list may be:
\begin{itemize}
 \item a binary erosion/dilation based on FFT;
 \item a set of deconvolution filters;
 \item a more memory efficient implementation of the convolution filter using the Overlap-Save algorithm.
\end{itemize}

>>>>

% 
% NOTE: This is the last number of the "handle" URL that 
% The Insight Journal assigns to your paper as part of the
% submission process. Please replace the number "1338" with
% the actual handle number that you get assigned.
%
\newcommand{\IJhandlerIDnumber}{}

@@

\begin{document}

%
% Add hyperlink to the web location and license of the paper.
% The argument of this command is the handler identifier given
% by the Insight Journal to this paper.
% 
\IJhandlefooter{\IJhandlerIDnumber}

\maketitle

@@

\IJhandlenote{\IJhandlerIDnumber}

@@

@@

All the performance tests are run on a workstation with two Intel(R) Xeon(R) CPU X5570 at 2.93GHz
and 24 GB of RAM and running Mandriva Linux 2009.1.

@@

As visible in figure \ref{fig:perf-kernel}, the kernel size as a little effect on the execution time
for the FFT based convolution, but a quite big impact on the normal one.

\begin{figure}[htbp]
\centering
\includegraphics{perf2D}
\caption{The effect of the kernel size on the execution time. The input image size is $256 \times 256$.
\label{fig:perf-kernel}}
\end{figure}

The drop in performance at some times seems to be due to a not so good plan generated by FFTW with \verb$FFTW_ESTIMATE$.
Using FFTW's wisdom feature can almost remove that problem. An integration of wisdom in ITK's FFTW
filters will be proposed in a later contribution.

@@

Most of the time, and despite the usual advice to use a size which is a power of two,
those sizes perform significantly better than the closest sizes which are a power of two, probably
because the performance gain per pixel doesn't compensate the higher number of pixels in the image.

@@

As visible in figure \ref{fig:perf-kernel}, the FFT based convolution and the normal one are scaling
quite well with the number of threads.

\begin{figure}[htbp]
\centering
\includegraphics{perf_threads}
\caption{The effect of the number of threads on the execution time. The input image size is
$371 \times 371 \times 34$, and the kernel size is two.
\label{fig:perf-kernel}}
\end{figure}

\section{Wrapping}

All the new filters have been wrapped using WrapITK.

\section{Development version}

A development version is available in a darcs repository at
\url{http://mima2.jouy.inra.fr/darcs/contrib-itk/fftconv/}.

@@

Some other work may be done on this basis -- an non exhaustive list may be:
\begin{itemize}
 \item a binary erosion/dilation based on FFT;
 \item a set of deconvolution filters;
 \item a more memory efficient implementation of the convolution filter using the Overlap-Save algorithm;
 \item let the user create a plan with \verb$FFTW_PATIENT$, \verb$FFTW_MEASURE$, or
       \verb$FFTW_EXHAUSTIVE$, so that FFTW can create better plan with \verb$FFTW_ESTIMATE$ in a
       later run;
 \item integrate FFTW's wisdom to improve FFT performance in ITK.
\end{itemize}
